```{r}
setwd("/Users/dongheelee/Desktop/pixmap_project")
install.packages("pixmap")
library(pixmap)
```

For each face, we get four views. 
Each row of the "all_faces" matrix represents the data for a version of a particular face.

```{r}
directories = dir(path="CroppedYale/",all.files=FALSE)
pics = 1:length(directories)
views = c('P00A+000E+00', 'P00A-005E+10','P00A-005E-10','P00A-010E+00')
all_faces = vector()
for(i in 1:length(pics)){
  
  add_matrix <- vector()
  for(j in 1:length(views)){
    file <- sprintf("CroppedYale/%s/%s_%s.pgm", directories[pics[i]], directories[pics[i]], views[j])

    face <- read.pnm(file) #get the face
    
    face_matrix <- getChannels(face)
    face_vector <- as.vector(face_matrix)
    
    add_matrix <- rbind(add_matrix, face_vector)
  }
  
  all_faces <- rbind(all_faces, add_matrix)
  dim(all_faces)
}
```


```{r}
mean_face <- colMeans(all_faces) #gets average pixel value across all of the faces (we get 32256 values in total), resulting a single row of values that represents the "mean face"

mean_face_image <- pixmapGrey(matrix(mean_face, nrow=192, ncol=168)) #this creates a bitmapped image 

#centers column by column

plot(mean_face_image)

centered_all_faces <- all_faces
for(i in 1:length(mean_face)){
  centered_all_faces[, i] <- centered_all_faces[, i]-mean_face[i]
}

plot(centered_all_faces)
#below centers row by row  

#center the mean_face_mat, will have dimesions 152 by 32256
#mean_face_centered = mat.or.vec(dim(all_faces)[1],dim(all_faces)[2])

#for each row (photo), subtract off the mean pix from the corresponding pix in the photo
#for(i in 1:dim(all_faces)[1]){
  #mean_face_centered[i,] <- all_faces[i,]-mean_face
#}
```

```{r}
```

I am applying pcr analysis on the image (mean-centered) matrix.
We have, in total, 152 principal components, each representing a photo that is called an eigenface. 
```{r}
eigenfaces <- prcomp(centered_all_faces)
```

Here are the first 9 eigenfaces in a 3-by-3 matrix. 
Each eigenface brings out particular features of the face. 
```{r}
row_of_images <- vector()
three_by_three <- vector()

for(i in 1:9){
  
  eigenface_matrix <- matrix(eigenfaces$rotation[,i], nrow=192, ncol=168) #192 by 168 is the dimension for a single face matrix, eigenfaces$rotation = column of eigenfaces 

  
  row_of_images <- cbind(row_of_images, eigenface_matrix)
  
  if (i %% 3 == 0){
    
    three_by_three <- rbind(three_by_three, row_of_images)
    row_of_images <- c()
  }
  
}
a <- pixmapGrey(three_by_three)
plot(a)
```

Plot of the number of components on the x-axis against the proportion of the variance explained on the y-axis:

```{r}
proportion_of_variance <- cumsum((eigenfaces$sdev)^2)*100/(sum((eigenfaces$sdev)^2))

plot(1:length(proportion_of_variance), proportion_of_variance, type="l", col="blue", xlab="Principal Component", ylab="Proportion of Variance")
```

I am using the eigenfaces to reconstruct the photo yaleB05_P00A+010E+00.pgm (this is the 4th pic in the 5th directory (overall, this is 4*5 = 20th pic).

I start at the mean face, and one by one, I add in an eigenface at a time until reaching 24 eigenfaces.   

 
```{r}
start <- mat.or.vec(nr=length(mean_face), nc = 1)
eigenface <- start
for(i in 1:24){
  #the length of a column of eigenfaces$x is 152 
  #each column of eigenfaces$rotation represents a face
  
  eigenface <- eigenface + eigenfaces$x[20, i]*eigenfaces$rotation[, i]
}

eigenface_matrix <- matrix(eigenface, nrow=192, ncol=168)
image <- pixmapGrey(eigenface_matrix)

plot(image)
```